### 2015-02-19 doc/proposal/proposal.tex
Initial draft for proposal

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amssymb}

\usepackage{hyperref}  % for urls and hyperlinks

\setlength{\textwidth}{6.2in}
\setlength{\oddsidemargin}{0.3in}
\setlength{\evensidemargin}{0in}
\setlength{\textheight}{8.7in}
\setlength{\voffset}{-.7in}
\setlength{\headsep}{26pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

% input some useful macros from RJLmacros.tex:
\input{../macros/RJLmacros}

\begin{document}
\section{Description}
We are planing to model the interaction of a bore and a free-standing coastal structure with Clawpack in 2D. In a first step we plan to model the wave tank followed by placing the coastal structure, a square column in the domain. We would like to compare the results with experimental results by Halld\'or \'Arnason.

\end{document}